---
title: "hw02-Jenny-Huang"
author: "Jenny Huang"
date: "9/27/2017"
output: github_document
---
```{r}
library(readr)
setwd("~/Desktop/stat133/stat133-hws-fall17/hw02")
```
## 2) Import the data in R  

```{r}
dat <- read.csv("./data/nba2017-player-statistics.csv", header = TRUE,
                  colClasses =c("Player" = "character",
                                "Team" = "character",
                                "Position" = "factor",
                                "Experience" = "character",
                                "Salary" = "real"
                                ))
str(dat)
```

```{r}
# importing data with read_csv()
read_csv("./data/nba2017-player-statistics.csv",
                  col_types = list(.default = col_integer(),
                                   Player = col_character(), 
                                   Team = col_character(), 
                                   Experience = col_character(),
                                   Position = col_factor(c("C","PF","PG","SF","SG")),
                                   Salary = col_double()))
str(read_csv("./data/nba2017-player-statistics.csv",
                  col_types = list(.default = col_integer(),
                                   Player = col_character(), 
                                   Team = col_character(), 
                                   Experience = col_character(),
                                   Position = col_factor(c("C","PF","PG","SF","SG")),
                                   Salary = col_double())))
```

## 3) Right after importing the data  

```{r}
#converting 'experience' into integers and replacing "R" with 1.
dat$Experience <- as.integer(replace(dat$Experience,grep("R",dat$Experience),"0"))
```

## 4) Performance of players  

```{r}
dat$Missed_FG <- dat$FGA - dat$FGM
dat$Missed_FG
dat$Missed_FT <- dat$FTA - dat$FTM
dat$Missed_FT
dat$PTS <- 3 * dat$Points3 + 2 * dat$Points2 + dat$FTM 
dat$PTS
dat$REB <- dat$OREB + dat$DREB
dat$REB
dat$MPG <- dat$MIN/dat$GP
dat$MPG
dat$EFF <- (dat$PTS + dat$REB + dat$AST + dat$STL + dat$BLK - dat$Missed_FG - dat$Missed_FT -dat$TO) / dat$GP
dat$EFF

summary(dat$EFF)
```

#### histogram
```{r}
hist(dat$EFF, xlab = "EFF", main = "Histogram of Efficiency(EFF)", col = "light grey")
```

#### top 10 players by EFF
```{r}
dat[order(-dat$EFF),][(1:10),c("Player","Team","Salary","EFF")]
```

#### Players with negative EFF
```{r}
dat[dat$EFF<0, "Player"]
```

#### correlation coefficients between EFF and other variables
```{r}
#correlation coefficients
cor(dat$EFF, dat$PTS)
cor(dat$EFF, dat$REB)
cor(dat$EFF, dat$STL)
cor(dat$EFF, dat$AST)
cor(dat$EFF, dat$BLK)
-cor(dat$EFF, dat$Missed_FT)
-cor(dat$EFF, dat$Missed_FG)
-cor(dat$EFF, dat$TO)

#correlation data frame
correlations <- data.frame(player_stats = c("PTS","REB","STL","AST","BLK","Missed_FT","Missed_FG","TO"),
cor = c(cor(dat$EFF, dat$PTS),
cor(dat$EFF, dat$REB),
cor(dat$EFF, dat$STL),
cor(dat$EFF, dat$AST),
cor(dat$EFF, dat$BLK),
-cor(dat$EFF, dat$Missed_FT),
-cor(dat$EFF, dat$Missed_FG),
-cor(dat$EFF, dat$TO)))

#display the computed correlations in descending order
descending_cor <- correlations[order(-correlations$cor),]

#creating barplot

barplot(descending_cor$cor, ylim = c(-1,1), 
        main = "Correlations between Player Stats and EFF", 
        col = c("orange","grey")[(correlations$cor>0) +1],
        names.arg = correlations$player_stats,
        cex.names = 0.5)
abline(h=0.0)
```
## 5) Efficiency and Salary  

```{r}
plot(dat$EFF,dat$Salary,
     xlab = "Efficiency", ylab = "Salary",
    col = "grey")
lines(lowess(dat$EFF,dat$Salary, f= 0.7), col = "red", lwd = 3)
#correlation between efficiency and salary
cor(dat$EFF,dat$Salary)
```
The correlation between EFF and Salary is 0.655624, which shows that there's a positive
relationship between EFF and Salary, that as EFF increases, salary goes up for players
in general. 

```{r}
#players with MPG of 20 or more minutes
players2 <- dat[dat$MPG >= 20,]
plot(players2$EFF,players2$Salary,
     xlab = "Efficiency", ylab = "Salary",
    col = "grey")
lines(lowess(players2$EFF,players2$Salary, f= 0.7), col = "red", lwd = 3)
#correlation between efficiency and salary
cor(players2$EFF,players2$Salary)
```
The correlation between EFF and Salary for players with minutes/game that's greater
than 20 is 0.5367, which means that there's a positive relationship between the two
variables, although the correlation is weaker than that of all players. This is probably
caused by the fact that a lot of experienced players don't play as often as they used
to nowadays so they have a basis salary that's lower, but their average minutes/game 
is still a lot higher than that of relatively new players or rookies. 

## 6) Comments and Reflections  

+ Creating the barplot with different features was hard.
+ Finding the correlations was pretty easy.
+ Yes. I got help from friends with read.csv because I had trouble locating the right 
repository.
+ It took around 3 hours.
+ Plotting barplot was the most time consuming part.